Introduction to the Tidyverse

```{r}
install.packages("gapminder")
install.packages("dplyr")

library("gapminder")
library("dplyr")

gapminder
```
Filtering for one year
```{r}

gapminder %>%
  filter(year==1957)

```
Filtering for one country and one year
```{r}
gapminder %>%
  filter(year==1957, country == "China")
```

Arranging observations by life expectancy
```{r}
gapminder %>%
  arrange(lifeExp)

#Sort in descending order of lifeExp

gapminder %>%
  arrange(desc(lifeExp))
```
Filtering and arranging

```{r}
gapminder %>%
  filter(year == 1957) %>% 
  arrange(desc(pop))
```

Using mutate to change or create a column

```{r}
# Use mutate to change lifeExp to be in months

gapminder %>%
  mutate(lifeExpMonths = 12 * lifeExp)
```
Combining filter, mutate and arrange

```{r}
gapminder %>%
  filter(year == 2007) %>% 
  mutate(lifeExpMonths = 12*lifeExp)
```


Visualizing with ggplot2

```{r}

library(ggplot2)

gapminder_1952 <- gapminder %>%
  filter(year == 1952)


ggplot(gapminder_1952, aes(x=pop, y=gdpPercap)) + 
  geom_point()

```
```{r}
ggplot(gapminder_1952, aes(x=pop, y=lifeExp)) + 
  geom_point()

```


Putting the a-axis on a log scale

```{r}

gapminder_1952 <- gapminder %>%
  filter(year == 1952)


ggplot(gapminder_1952, aes(x=pop, y=lifeExp)) + 
  geom_point() + 
  scale_x_log10()

```

Putting the x-and and y-axes on a log scale

```{r}

ggplot(gapminder_1952, aes(x=pop, y=gdpPercap))+
  geom_point()
  scale_x_log10()+
  scale_y_log10()

```

Adding color to a scatter plot

```{r}

gapminder_1952 <- gapminder %>%
  filter(year == 1952)

ggplot(gapminder_1952, aes(x=pop, y=lifeExp, color = continent)) + 
  geom_point() +
  scale_x_log10()

```

Adding size and color to a plot

```{r}
ggplot(gapminder_1952, aes(x=pop, y=lifeExp, color = continent, size = gdpPercap)) + 
  geom_point() +
  scale_x_log10()
```

Faceting

Creating a subgraph for each continent

```{r}
gapminder_1952

ggplot(gapminder_1952, aes(x=pop, y=lifeExp)) + 
  geom_point() +
  scale_x_log10() +
  facet_wrap(~ continent)
  

```

faceting by year

```{r}

ggplot(gapminder, aes(x=gdpPercap, y=lifeExp, color= continent, size=pop)) + 
  geom_point() +
  scale_x_log10() +
  facet_wrap(~ year)
  

```

